---
title: "JohnsHopkins"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quizz 1 - Linear Regression

Question 1.

```{r }
x <- c(0.18, -1.54, 0.42, 0.95)
w <- c(2, 1, 3, 1)
u <- c(1.077, 0.3, 0.1471, 0.0025)

for (i in u)
{
  resultat <- w[1]*(x[1]-i)^2 + w[2]*(x[2]-i)^2 + w[3]*(x[3]-i)^2
  print(i)
  print(resultat)
}



```

Question 2.  

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
model <- lm(y~x+0) # Adding zeor to go through the origin
summary(model)
```

Question 3

```{r}
data(mtcars)
```

```{r}
library(ggplot2)
ggplot(mtcars) + 
    geom_point(mapping=aes(x=wt,y=mpg))
```

Strong correlation & negative.  

```{r}
mpg_model <- lm(mtcars$mpg~mtcars$wt)
summary(mpg_model)
```

Question 4   
slope = sy / sx * R  
The standard deviation of the predictor is one half that of the outcome  
sx = 0.5 sy
sy / (0.5 * sy) = 2  
R = 0.5

```{r}
slope <- 1
```


Question 5  
Slope = sy / sx * R  
sy = sx = 1 - the results were normalized mean 0; variance 1  
R = 0.4  
slope = 0.4  

```{r}
score2 <- 1.5 * 0.4
score2
```

Question 6  
Data normalization  
We get mean = 0 and sd = 1
This helps to get rid of shifting or scaling effects.  
Zi = (Xi - X_bar)/sd 


```{r}
x <- c(8.58, 10.46, 9.01, 9.64, 8.86)
z1 <- (x[1]-mean(x))/sd(x)
z1
```

Question 7  
Fitting a another linear model  

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
my_model <- lm(y~x)
summary(my_model)

```


Question 8  
Remember that : intercept = y_bar - slope * x_bar  
so, if x_bar = y_bar = 0 => The intercept is zero.  


question 9




```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
for (i in x)
{
  resultat <- sum((x-i)^2)
  print(i)
  print(resultat)
}
```

Question 10  
beta1 = sy/sx * R
gamma 1 = sx / sy * R

beta1/gamma1 = (sy/sx)/(sx/sy) = sy² / sx ² = varY / varX


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
