---
title: "JH - M07 - Course Notes"
author: "shostiou"
date: "13/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploration of Galton's data

reshape est une extension développée par Hadley Wickham. Elle permet de passer facilement d'un format de données «long» (une ligne par individu et par variable) à un format «large» (une seule ligne par individu). Elle est notamment utile pour mettre en forme les données avant utilisation des fonctions graphiques de ggplot2.  

galton :  
ligne = 1 colonne child / 1 colonne parent  
long :  
1 ligne = 1 donnée - 1 colonne variable / 1 colonne valeur  

```{r}
library(UsingR)
library(reshape)
library(ggplot2)
```



```{r explo}

# the melt function is part of the reshape package
long <- melt(galton)
# plotting the data
g <- ggplot(long, aes(x=value, fill = variable))+
    geom_histogram(colour = "black", binwidth=1)+
    facet_grid(. ~ variable)
g
```

## Research the Middle via Least squares

```{r}
library(manipulate)
```

```{r}
myHist <- function(mu){
# Computation of mean squarred error
mse <- mean((galton$child - mu)^2)
# plotting mean value
g <- ggplot(galton, aes(x = child)) + geom_histogram(fill = "salmon", colour = "black", binwidth=1)
g <- g + geom_vline(xintercept = mu, size = 3)
g <- g + ggtitle(paste("mu = ", mu, ", MSE = ", round(mse, 2), sep = ""))
g
}
manipulate(myHist(mu), mu = slider(62, 74, step = 0.5))
```

It seems we minimize the value of MSE at mu = 68.  

```{r}
mean(galton$child)
```

The empirical mean minimizes the function.  
Scatter plot

```{r}
ggplot(galton, aes(x=parent,y=child))+geom_jitter() + geom_smooth(method="lm")
```

The jitter function is used because a lot of points are overlapped.  
The geom_smooth can be used (with proper lm) to trace the lm approximation.  



## Fitting Galton's data


Fitting Galton's data with linear regression  

```{r}
x <- galton$parent
y <- galton$child
# correlation coefficient
R <- cor(x,y)
sx <- sd(x)
sy <- sd(y)
# slope
beta1 <- sy/sx*R
# using mean values as the line goes through these point.
beta0 <- mean(y) - beta1*mean(x)  
# Printing
paste("intercept: ",beta0,"slope : ",beta1)
# Comparing result to the lm fit function
my_model <- lm(y~x)
summary(my_model)
```


## Example of Regression - diamond price  

Let’s analyze the diamond data set from the UsingR package. The data is diamond prices (in Singapore
dollars) and diamond weight in carats. Carats are a standard measure of diamond mass, 0.2 grams.
To get the data use library(UsingR); data(diamond)  

```{r}
data(diamond)
```

```{r}
ggplot(data=diamond, aes(x=carat, y=price))+geom_point() + geom_smooth(method="lm")
```

Training the model

```{r}
my_model <- lm(price~carat,data=diamond)
summary(my_model)
```

```{r}
coef(my_model)

```

Interpretation of the residuals  

```{r}
# The resid function records residuals - model fit
my_resid <- resid(my_model)
# Preduction of the model
my_pred <- predict(my_model)

# Controlling constant variance
plot(my_resid)
hist(m_resid)
```






