---
title: "JH - M07 - W02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 01

Consider the following data with x as the predictor and y as as the outcome.  
Give a P-value for the two sided hypothesis test of whether β1 \beta_1 β1.  
Let's train the linear model and check the p-value corresponding to H0

```{r cars}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
my_model <- lm(y~x)
summary(my_model)

```
As we have a p-value greater than 0.05, we fail to reject the null Hypothesis.  


## Question 02

For the previous problem, let's consider the sd of the residuals.  
This is equal to 0.223.  
  
## Question 3  
In the mtcars\verb|mtcars|mtcars data set, fit a linear regression model of weight (predictor) on mpg (outcome). Get a 95% confidence interval for the expected mpg at the average weight. What is the lower endpoint?


```{r}
data(mtcars)
```

```{r}
library(ggplot2)
ggplot(mtcars) + 
    geom_point(mapping=aes(x=wt,y=mpg))
```

```{r}
# Getting the mean value of wt
wt_mean <- mean(mtcars$wt)
# fitting the linear model
mod_mpg <- mtcars$mpg
mod_wt <- mtcars$wt
mpg_model <- lm(mod_mpg~mod_wt)
# displaying summary
summary(mpg_model)

```
Prediction with a 95% confidence interval a mean value of wt

```{r}
wt_mean <- mean(mtcars$wt)
my_in <- data.frame(mod_wt = wt_mean)
predict(mpg_model,my_in, interval="confidence")
```
The lower end point is 18.99  
  


## Question 5  

Consider again the mtcars\verb|mtcars|mtcars data set and a linear regression model with mpg as predicted by weight (1,000 lbs). A new car is coming weighing 3000 pounds. Construct a 95% prediction interval for its mpg. What is the upper endpoint?


```{r}
new_car<- 3000/1000
my_in <- data.frame(mod_wt = new_car)
predict(mpg_model,my_in, interval="prediction")
```

The confidence interval reflects the uncertainty around the mean predictions. To display the 95% confidence intervals around the mean the predictions, specify the option interval = "confidence":

The prediction interval gives uncertainty around a single value. In the same way, as the confidence intervals, the prediction intervals can be computed as follow

## Question 6  

Consider again the mtcars\verb|mtcars|mtcars data set and a linear regression model with mpg as predicted by weight (in 1,000 lbs). A “short” ton is defined as 2,000 lbs. Construct a 95% confidence interval for the expected change in mpg per 1 short ton increase in weight. Give the lower endpoint.

```{r}
new_car<- 2000/1000
my_in <- data.frame(mod_wt = new_car)
predict(mpg_model,my_in, interval="confidence")
confint(mpg_model)
```
The answer will be

```{r}
 -6.486308 *2
```
One short ton increase is equivalent to x2 units.  


## Question 9

```{r}
# Model with just the intercept
sum((mtcars$mpg-(mtcars$wt+37.2851))^2)
# Complete model
sum(mpg_model$residuals^2)
```



