---
title: "Johns Hopkins Coursera - Statistical Inference - Project Part 2"
author: "shostiou"
date: "06/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 02 - Basic Inferential Data Analysis

This second part of the project is dedicated to the analysis of the ToothGrowth dataset (embedded in the R package).  

### Data upload & basic exploration

Let's start by loading the dataset and by doing basic explorations.  
Refering to the help associated to the data, this dataset is describe as : **"The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC)."**  

Content :   
60 observations on 3 variables.
[,1]	len	numeric	Tooth length
[,2]	supp	factor	Supplement type (VC or OJ).
[,3]	dose	numeric	Dose in milligrams/day

```{r data_load}
# Loading the dataset
data(ToothGrowth)  
# Getting general overview of the data  
str(ToothGrowth)  
# Getting summary statistics.
summary(ToothGrowth)  
# Let's visualize the distribution of the numerical values
hist(ToothGrowth$len, main = "distribution of Tooth length")  
hist(ToothGrowth$dose, main = "distribution of Vitamin doses")

```
We can observe that the data contains 2 groups of observations. Each of those groups is made of 30 observation (Orange Juice / Ascorbic Acid).  
Even though the vitamin doses is numerical data, the values observed in the dataset is limited to 3 levels : 0.5; 1 ; 2. We can consider this variable as a discrete (not coninuous).  

Let's pursue the exploration by comparing the distribution of Tooth length based on the delivery method used (orange Juice - OJ ; ascorbid acid - VC)


```{r hist_by_del_method}
# We call use ggplot2 to show the distributions  
library(ggplot2)
# Building the histograms based on supp variable  
ggplot(ToothGrowth, aes(x = len)) +
  geom_histogram(aes(color = supp, fill = supp), 
                position = "identity", alpha = 0.4) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))

```
The observation of the 2 overlapped distributions doesn't provide a key evidence between Tooth lengths & the delivery method used during the experiment.


### Comparing the means

Let's compare the means of the tooth length distribution based on the delivery method.

```{r sample_means}
# Let's call the dplyr package
library(dplyr)
# means & standard error computation  
ToothGrowth %>% group_by(supp) %>% summarize(mean_len = mean(len))

```

We can observe that mean values of the 2 distributions (tooth length with orange juice OJ / tooth length with ascorbic acis CV) differs.  

### Hypothesis Test

Now to determine is this difference is statistically pertinent, we will use a Hypothesis test :
H0 : xbar_len_OJ - xbar_len_CV = 0
HA : xbar_len_OJ -  xbar_len_CV != 0  




