---
title: "JH_capstone_W06"
author: "shostiou"
date: "06/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this step, the purpose is to finalize the prediction algorithm and implement a Shiny App to build a prediction of the next 3 words.  
A stupid backoff approach will be used : starting from 4-gram to 2-gram

References :  
Since good estimates can be made based on smaller models, it is more practical to use bi- or trigram models. This idea that a future event (in this case, the next word) can be predicted using a relatively short history (for the example, one or two words) is called a Markov assumption.  

https://cs.stanford.edu/people/eroberts/courses/soco/projects/2004-05/nlp/techniques_word.html


# Libraries

```{r, message=FALSE,warning=FALSE}
library(R.utils)
library(RSQLite)
library(tictoc)
library(stringr)
library(tm)
library(dplyr)
library(tidytext)
library(tidyr)
```




# Algorithm prototype

String used as an input argument

```{r}
# input argument
my_string <- tolower(removePunctuation("i"))

```

This approach relies on a simplified Markov chain rule which embeds 4-grams to 2-grams.  
As a stupid backoff stategy, we start by checking if a prediction corresponding to 4-grams exists. If, not a new search based on 3-grams is built and so on.


```{r}
pred_next_word<- function(mystring){

    # tic toc functions used to evaluate performance (response time) 
    # tic("predicting a word version 2")
    
    # connecting to database  
    mydb <- dbConnect(RSQLite::SQLite(), "ngrams_db_prob.sqlite")
    
    # Counting the number of words in the string
    nb_words <- str_count(my_string,"\\S+")
    word_1=''
    word_2=''
    word_3=''
    
    # init of the variable used to check if a result exists
    is_result <- 0
    
    # Looking for 4-grams
    if (nb_words >=3)
      {
      # Getting the words  
      word_1 = word(my_string,-1)
      word_2 = word(my_string,-2)
      word_3 = word(my_string,-3)
      
      # Building query to the DB  
      query4g <- paste0("SELECT word4, prob_n FROM fourgram_prob WHERE word3 LIKE '", word_1, 
                     "' AND word2 LIKE '",word_2,
                     "' AND word1 LIKE '",word_3,"' ORDER BY prob_n DESC LIMIT 3")
      result <- dbGetQuery(mydb, query4g)
      
      # Looking for an empty result
      is_result <- nrow(result)
      # if (is_result != 0){print(result)} 
      
      } 
    
    
    # Looking for 3-grams
    if ( ((nb_words ==2)| (is_result == 0)) & !(nb_words==1))
      {
      # Getting the words  
      word_1 = word(my_string,-1)
      word_2 = word(my_string,-2)
    
      # Building query to the DB  
      query3g <- paste0("SELECT word3,prob_n FROM trigram_prob WHERE word2 LIKE '", word_1, 
                     "' AND word1 LIKE '",word_2,"' ORDER BY prob_n DESCLIMIT 3")
      result <- dbGetQuery(mydb, query3g)
      
      # Looking for an empty result
      is_result <- nrow(result)
      # if (is_result != 0){print(result)} 
      
      } 
    
    # Looking for 2-grams
    if ((nb_words ==1)| (is_result == 0) )
      {
      # Getting the words  
      word_1 = word(my_string,-1)
    
      # Building query to the DB  
      query2g <- paste0("SELECT word2,prob_n FROM bigram_prob WHERE word1 LIKE '", word_1,"' ORDER BY prob_n DESC LIMIT 3")
      result <- dbGetQuery(mydb, query2g)
      
      # Looking for an empty result
      is_result <- nrow(result)
      # if (is_result != 0){print(result)} 
      
      } 
    
    # disconnecting
      if (is_result != 0)
        {
        result1 <- as.character(unlist(as.list(result)[1])[1])
        prob1 <- as.numeric(unlist(as.list(result)[2])[1])
        result2 <- as.character(unlist(as.list(result)[1])[2])
        prob2 <- as.numeric(unlist(as.list(result)[2])[2])
        result3 <- as.character(unlist(as.list(result)[1])[3])
        prob3 <- as.numeric(unlist(as.list(result)[2])[3])
        } 
    
    dbDisconnect(mydb)
    # toc()
    
    #returning result
    # print(result)
    return(list(result1,prob1,result2,prob2,result3,prob3))
    
    }

```

```{r}
result1 <- unlist(pred_next_word(my_string))[1] 
result1
```
```{r}
result2
```












```{r}
my_string <- tolower(removePunctuation("you know i love"))
# number of words in the string  
nb_words <- str_count(my_string,"\\S+")
word_2=''
word_3=''
word_4=''
```



```{r}
tic('version2')
mydb <- dbConnect(RSQLite::SQLite(), "my-db.sqlite")

# Extracting last words from the string
word_1 = word(my_string,-1)
query2g <- paste0("SELECT * FROM bigram_prob WHERE word1 LIKE '", word_1,"' LIMIT 3")
dbGetQuery(mydb, query2g)

if (nb_words >=2){word_2 = word(my_string,-2)
query3g <- paste0("SELECT * FROM trigram_prob WHERE word2 LIKE '", word_1, 
                 "' AND word1 LIKE '",word_2,"' LIMIT 3")
dbGetQuery(mydb, query3g)} 


if (nb_words >=3){word_3 = word(my_string,-3)
query4g <- paste0("SELECT * FROM fourgram_prob WHERE word3 LIKE '", word_1, 
                 "' AND word2 LIKE '",word_2,
                 "' AND word1 LIKE '",word_3,"' LIMIT 3")
dbGetQuery(mydb, query4g)}


if (nb_words >=4){word_4 = word(my_string,-4)
query5g <- paste0("SELECT * FROM fivegram_prob WHERE word4 LIKE '", word_1, 
                 "' AND word3 LIKE '",word_2,
                 "' AND word2 LIKE '",word_3,
                 "' AND word1 LIKE '",word_4,"' LIMIT 3")
# dbGetQuery(mydb, query5g)
} 

dbDisconnect(mydb)
toc()
```

Getting data from the n-grams

```{r}







```

Send the request to the database  

```{r}
mydb <- dbConnect(RSQLite::SQLite(), "my-db.sqlite")





```
