---
title: "JH_captone_w03"
author: "shostiou"
date: "25/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries  

```{r, echo = FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(ggplot2)
library(tidytext)
library(readtext)
library(corpus)
library(R.utils)
library(tm)
library(stringr)
```

Importing the unigrams, bigrams and trigrams defined during the previous week.


load("en_US_token.Rda")
load("en_US_bigram.Rda")
load("en_US_trigram.Rda")
load("en_US_fourgram.Rda")
load("en_US_fivegram.Rda")


##  Splitting the words into distinct columns

In order to setup of prediction environment, we need to be split our bigram data set into 2 columns : word1 and word 2 and split the data into 3 columns for the trigram.  

```{r, cache=TRUE}
# load("en_US_bigram.Rda")
en_US_bigram <- en_US_bigram %>% 
  separate(bigram, c("word1","word2"), sep =" ")
save(en_US_bigram,file="en_US_bigram_split.Rda")
rm(en_US_bigram)
```


```{r, cache=TRUE}
load("en_US_trigram.Rda")
en_US_trigram <- en_US_trigram %>% 
separate(trigram, c("word1","word2","word3"), sep =" ")
save(en_US_trigram,file="en_US_trigram_split.Rda")
rm(en_US_trigram)
```

```{r, cache=TRUE}
load("en_US_fourgram.Rda")
en_US_fourgram <- en_US_fourgram %>% 
  separate(fourgram, c("word1","word2","word3","word4"), sep =" ")
save(en_US_fourgram,file="en_US_fourgram_split.Rda")
rm(en_US_fourgram)
```


```{r, cache=TRUE}
load("en_US_fivegram.Rda")
en_US_fivegram <- en_US_fivegram %>% 
  separate(fivegram, c("word1","word2","word3","word4","word5"), sep =" ")
save(en_US_fivegram,file="en_US_fivegram_split.Rda")
rm(en_US_fivegram)
```


# Defining probabilities for predictions 

Our prediction algorithm will be using probabilities to make a prediction of the next word. We can think about the following approach :  
- taking word1 as an input argument, which probability is associated to word 2.  
- the same approach will be extended to tri-grams, but this time, the probability to be taken into consideration will be defined by considering word1 & word2 as input arguments.  

Defining probabilities for bigrams

```{r}
load('en_US_bigram_split.Rda')
en_US_bigram_prob <- en_US_bigram %>% group_by(word1) %>% mutate(prob_word2 = n / sum(n)) %>% arrange(desc(word1)) %>% select(word1,word2,prob_word2)
save(en_US_bigram_prob,file='en_US_bigram_prob.Rda')
```


Defining probabilities for trigrams  

```{r}
load('en_US_trigram_split.Rda')
en_US_trigram_prob <- en_US_trigram %>% group_by(word1,word2) %>% mutate(prob_word3 = n / sum(n)) %>% arrange(desc(word1,word2)) %>% select(word1,word2,word3,prob_word3)
save(en_US_trigram_prob,file='en_US_trigram_prob.Rda')
```


Defining probabilities for fourgrams  

```{r}
load('en_US_fourgram_split.Rda')
en_US_fourgram_prob <- en_US_fourgram %>% group_by(word1,word2,word3) %>% mutate(prob_word4 = n / sum(n)) %>% arrange(desc(word1,word2,word3)) %>% select(word1,word2,word3,word4,prob_word4)
save(en_US_fourgram_prob,file='en_US_fourgram_prob.Rda')
```


Defining probabilities for fivegrams  

```{r}
load('en_US_fivegram_split.Rda')
en_US_fivegram_prob <- en_US_fivegram %>% group_by(word1,word2,word3,word4) %>% mutate(prob_word5 = n / sum(n)) %>% arrange(desc(word1,word2,word3,word4)) %>% select(word1,word2,word3,word4,word5,prob_word5)
save(en_US_fivegram_prob,file='en_US_fivegram_prob.Rda')
```



# Proof of concept

Googling and reading discussions in the course forum helped me to clarify the strategy to be applied to this project.  
Many mentors were underlying the importance of "keeping the model" simple.  
This means that applying highly sophisticated techniques will not have a significant effect on containing the prediction error of the model.  
That's why I decided to stick to my initial approach : prediction based on probabilities related to n-grams.  


##  Using SQLite to speedup performances - Building tables

Tibbles are too small to perform search requests in large volume of data.
Some web researches highlighted that using a SQLite database could contribute to significantly improve the performance of data retrieval.

```{r}
library(RSQLite)
```

Creating the connection to the database.

```{r}
mydb <- dbConnect(RSQLite::SQLite(), "my-db.sqlite")
```

Building the five grams table

```{r}
load('en_US_fivegram_prob.Rda')
dbWriteTable(mydb, "fivegram_prob", ungroup(en_US_fivegram_prob))

```

Building the four grams table

```{r}
load('en_US_fourgram_prob.Rda')
dbWriteTable(mydb, "fourgram_prob", ungroup(en_US_fourgram_prob))
```


Building the three grams table

```{r}
load('en_US_trigram_prob.Rda')
dbWriteTable(mydb, "trigram_prob", ungroup(en_US_trigram_prob))
```

Building the bi grams table


```{r}
load('en_US_bigram_prob.Rda')
dbWriteTable(mydb, "bigram_prob", ungroup(en_US_bigram_prob))
```

Disconnecting the database

```{r}
dbDisconnect(mydb)
```



##  Doing some tests

```{r}
library(RSQLite)
library(tm)
library(stringr)
```

Creating the connection to the database.

```{r}
mydb <- dbConnect(RSQLite::SQLite(), "my-db.sqlite")
```

# Setting the words for prediction  

1) The guy in front of me just bought a pound of bacon, a bouquet, and a case of
=> beer - OK

2) You're the reason why I smile everyday. Can you follow me please? It would mean the
=> world - OK


3) Hey sunshine, can you follow me and make me the
=> happiest - OK

4) Very early observations on the Bills game: Offense still struggling but the
=< players - random guessing // Defense.


5) Go on a romantic date at the
=> beach - OK

6) Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my
=> way - OK

7) Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some
=> time - OK

8) After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little
=< random answer / fingers

9) Be grateful for the good times and keep the faith during the
=< bad - random guessing // OK for bad

If this isn't the cutest thing you've ever seen, then you must be

=< asleep - random guessing // insensitive - random guessing NOK



```{r}
my_string <- tolower(removePunctuation("Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his"))
# number of words in the string  
nb_words <- str_count(my_string,"\\S+")
word_2=''
word_3=''
word_4=''
# Extracting last words from the string
word_1 = word(my_string,-1)
if (nb_words >=2){word_2 = word(my_string,-2)} 
if (nb_words >=3){word_3 = word(my_string,-3)} 
if (nb_words >=4){word_4 = word(my_string,-4)} 
```

Getting data from the n-grams

```{r}
query5g <- paste0("SELECT * FROM fivegram_prob WHERE word4 LIKE '", word_1, 
                 "' AND word3 LIKE '",word_2,
                 "' AND word2 LIKE '",word_3,
                 "' AND word1 LIKE '",word_4,"' LIMIT 30")

query4g <- paste0("SELECT * FROM fourgram_prob WHERE word3 LIKE '", word_1, 
                 "' AND word2 LIKE '",word_2,
                 "' AND word1 LIKE '",word_3,"' LIMIT 30")

query3g <- paste0("SELECT * FROM trigram_prob WHERE word2 LIKE '", word_1, 
                 "' AND word1 LIKE '",word_2,"' LIMIT 30")

query2g <- paste0("SELECT * FROM bigram_prob WHERE word1 LIKE '", word_1,"' LIMIT 30")

```

Send the request to the database  

```{r}
dbGetQuery(mydb, query5g)
dbGetQuery(mydb, query4g)
dbGetQuery(mydb, query3g)
dbGetQuery(mydb, query2g)
```



word4 = world


```{r}
en_US_trigram_prob %>% filter(word1 == myt_word3,word2 ==myt_word4)
```

### Observations  

This is quite close to a "brute force" approach.  
The more data you get, the best predictions you get in return.  
This approach is quite similar to a KNN algorithm.  
The system works more or less like a memory : it restitutes what it has already seen !

The accuracy of the predictions will be improved by :  
- extending the sample size (currently limited to 5% of the overall datasets)








