---
title: "JH - M07 - W03 - Quiz 03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(ggplot2)
```




## QUESTION 1  

Consider the  mtcars\verb| mtcars| mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

```{r}
data("mtcars")
```

```{r}
my_model <- lm(mpg~wt+factor(cyl), data=mtcars)
summary(my_model)
```

This answer will be -6.07


## QUESTION 2  

Consider the  mtcars\verb| mtcars| mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?.  

```{r}
## Fitting a model without weigth
my_model2 <- lm (mpg~factor(cyl),data=mtcars)
summary(my_model2)

```

## QUESTION 3

Consider the mtcars\verb|mtcars|mtcars data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.






